NN_Accel_Heading

### 2024-03-13 Fish Data Analysis/NN_Accel_Heading.qmd
Added single fish to processing and graphs. Also made it an angle ring

## Loading in the Libraries for Everything
```{r}
library(MASS)
library(tidyverse)
library(ggplot2)
library(bbmle) #For ICtab
library(car)
library(ggpubr)
library(ggdist)
library(emmeans)
library(lme4)
library(viridis)
```

```{r}
rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}
round_any <- function(x, accuracy, f=round){f(x/ accuracy) * accuracy}
ang_mean <- function(x){rad2deg(atan2(mean(sin(deg2rad(x))),mean(cos(deg2rad(x)))))}

@@

fish_comp_data <- read.csv("Data/Fish_Comp_Values_3D.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"))

@@

min_dist <- 0
max_dist <- 0.75

fish_data_nn <- fish_comp_data %>% group_by(Fish1,Tailbeat_Num,Data_ID) %>%
                                    mutate(#nndDistance = min(Distance, na.rm = TRUE),
                                           nndDistance = nth_min(Distance, 1),
                                           nndDistance = ifelse(is.infinite(nndDistance), NA, nndDistance),
                                           all_nas = all(is.na(nndDistance))) %>%
                                    #filter(Distance == nndDistance | all_nas) %>%
                                    mutate(row_n = row_number()) %>%
                                    #filter(row_n == 1) %>%
                                    ungroup() %>%
                                    mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS) %>%
                                    filter(Distance >= min_dist & Distance <= max_dist)

fish_data_nn_normal <- fish_data_nn %>% group_by(Flow_Ablation_Darkness) %>%
                                        mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS - mean(Fish1_Yaw_Heading_PS)) %>%
                                        ungroup()

nn_heading_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Angle,pi/6)) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

nn_heading_dist_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Angle,pi/6), distance_bins = round_any(Distance,0.25)) %>%
                                                group_by(angle_bins,distance_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

nn_accel_sum <- fish_data_nn_normal %>% mutate(distance_bins = round_any(Relative_X,0.2)) %>%
                                                group_by(distance_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_accel = mean(Fish1_Accel, na.rm = T),
                                                          sd_accel = sd(Fish1_Accel, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

@@

heading_ps_angle_step <- stepAIC(heading_ps_angle_base,
                                   list(upper = ~ Angle*sin(Angle)*Ablation*Darkness, lower = ~1),
                                   direction = "both",
                                   trace = F)

@@

manual_model <- lm(Fish1_Yaw_Heading_PS ~ sin(Angle) + sin(Angle):Ablation + sin(Angle):Darkness + sin(Angle):Ablation:Darkness, 
                   data = fish_data_nn_model) 
Anova(manual_model)

#Raw data
ggplot(fish_data_nn %>% filter(Flow == "Still Water"),
       aes(x = Angle, y = rad2deg(Fish2_Yaw_Heading_PS), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_line(aes(y = predict(heading_ps_angle_step)), size = 1)+
  theme_classic() +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")

#Means graph
ggplot(nn_heading_sum %>% filter(Flow == "Still Water"),
       aes(x = angle_bins, y = rad2deg(mean_turn), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  #geom_line()+
  geom_line(aes(y = rad2deg(predict(manual_model,
                    newdata = nn_heading_sum %>% filter(Flow == "Still Water") %>% mutate(Angle = angle_bins)))), size = 1)+
  theme_classic() +
  ylim(-50,50) +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (degrees/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")+
  ggtitle(paste0("All Fish From ",min_dist," to ",max_dist))

test(emmeans(manual_model, pairwise ~Ablation*Darkness | Angle, at = list(Angle = c(pi/2))))

# ggplot(nn_heading_dist_sum %>% filter(Flow == "Still Water"),
#        aes(x = angle_bins, y = distance_bins, fill = mean_turn))+
#   facet_wrap(~ Ablation + Darkness)+
#   scale_fill_viridis(option="turbo",limits = c(-2,2))+
#   geom_tile()+
#   scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
#                      labels = round(unique(nn_heading_sum$angle_bins))) +
#   coord_polar(start = -pi / 2)+
#   theme_classic()

@@

```{r}
#par 1 = amp
#par 2 = freq
#par 3 = phase

min.RSS <- function(data, par) {
             with(data, sum(par[1]*sin(2*pi/par[2]*Angle+par[3]) - Fish1_Yaw_Heading_PS)^2)
}

result_F0LNDN <- optim(par = c(1, 1, 0), fn = min.RSS, data = fish_data_nn_normal %>% filter(Flow_Ablation_Darkness == "Still Water, No Ablation, Light"))
result_F0LYDN <- optim(par = c(1, 1, 0), fn = min.RSS, data = fish_data_nn_normal %>% filter(Flow_Ablation_Darkness == "Still Water, Ablated, Light"))
result_F0LNDY <- optim(par = c(1, 1, 0), fn = min.RSS, data = fish_data_nn_normal %>% filter(Flow_Ablation_Darkness == "Still Water, No Ablation, Dark"))
result_F0LYDY <- optim(par = c(1, 1, 0), fn = min.RSS, data = fish_data_nn_normal %>% filter(Flow_Ablation_Darkness == "Still Water, Ablated, Dark"))

result_F0LNDN$par
result_F0LYDN$par
result_F0LNDY$par
result_F0LYDY$par
```

accel_base_lm <- lm(Fish1_Accel ~ Flow_Ablation_Darkness, data = fish_data_nn %>% filter(Flow == "Flowing Water (2 BL/s)"))
accel_lin_lm <- lm(Fish1_Accel ~ Relative_X*Flow_Ablation_Darkness, data = fish_data_nn %>% filter(Flow == "Flowing Water (2 BL/s)"))

AIC(accel_base_lm,accel_lin_lm) %>% arrange(AIC)

Anova(accel_base_lm)

ggplot(fish_data_nn %>% filter(Flow == "Flowing Water (2 BL/s)"),
       aes(x = Relative_X, y = Fish1_Accel, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_smooth(se = F, method = glm, formula = y ~ x)+
  theme_classic() +
  xlab("distance from neighbour (BL)") +
  ylab("focal fish acceleration (BL/s^2)") +
  #ylim(-0.01, 0.01) +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  labs(color = "Sensory Conditions")

ggplot(nn_accel_sum %>% filter(Flow == "Flowing Water (2 BL/s)" & distance_bins < 3),
       aes(x = distance_bins, y = mean_accel, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  #geom_line(alpha = 0.75) +
  geom_smooth(se = F, method = glm, formula = y ~ x)+
  theme_classic() +
  xlab("distance from neighbour (BL)") +
  ylab("focal fish acceleration (BL/s^2)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  labs(color = "Sensory Conditions")

# directional_df <- fish_data_nn %>% mutate(round_X = round_any(Relative_X,0.5),
#                                                     round_Y = round_any(Relative_Y,0.5)) %>%
#                                                       group_by(round_X,round_Y,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
#                                                       summarise(mean_accel = mean(Fish1_Accel),
#                                                                 mean_turn = mean(Fish1_Yaw_Heading_PS),
#                                                                 n = n()) %>%
#                                                       ungroup()
# 
# ggplot(directional_df %>% filter(Flow == "Still Water" & n > 5 & mean_accel > -0.01 & mean_accel < 0.01),
#        aes(round_X, round_Y, fill= mean_accel)) + 
#   geom_tile()+
#   facet_wrap(~Flow_Ablation_Darkness)+
#   scale_fill_viridis(option="magma")+
#   coord_fixed()+
#   theme_classic()
# 
# 
# ggplot(directional_df %>% filter(Flow == "Still Water" & n > 5),
#        aes(round_X, round_Y, fill= mean_turn)) + 
#   geom_tile()+
#   facet_wrap(~Flow_Ablation_Darkness)+
#   scale_fill_viridis(option="magma")+
#   coord_fixed()+
#   theme_classic()

>>>>

## Loading in the Libraries for Everything

```{r}
library(MASS)
library(tidyverse)
library(ggplot2)
library(bbmle) #For ICtab
library(car)
library(ggpubr)
library(ggdist)
library(emmeans)
library(lme4)
library(viridis)
library(lme4)

```{r}
rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}
round_any <- function(x, accuracy, f=round){f(x/ accuracy) * accuracy}

deg_ang_mean <- function(x){rad2deg(atan2(mean(sin(deg2rad(x))),mean(cos(deg2rad(x)))))}
rad_ang_mean <- function(x){atan2(mean(sin(x)),mean(cos(x)))}

@@

single_comp_data <- read.csv("Data/Fish_Comp_Values_3D_Single.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "S", "Single","Jingle")) %>%
                           mutate(Darkness = ifelse(Darkness == "S", "Single","Jingle")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("Single"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Single"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                           # ### Used to remove every third point to reduce autocorrelation
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"),
                                  Date = paste(Year,Month,Day,Trial, sep = "_"),
                                  Fish_ID = paste(Year,Month,Day,Trial,Fish1, sep = "_"),
                                  Comp_ID = paste(Year,Month,Day,Trial,Fish1,Fish2, sep = "_")) %>%
                           group_by(Comp_ID) %>%
                           filter(row_number() %% 3 == 1) %>%
                           ungroup()

single_comp_data <- na.omit(single_comp_data)

fish_comp_data <- read.csv("Data/Fish_Comp_Values_3D.csv") %>%
                           mutate(Flow = ifelse(Flow == "0", "Still Water", "Flowing Water (2 BL/s)")) %>%
                           mutate(Ablation = ifelse(Ablation  == "N", "No Ablation", "Ablated")) %>%
                           mutate(Darkness = ifelse(Darkness == "N", "Light", "Dark")) %>%
                           mutate(Flow = factor(Flow), Ablation = factor(Ablation), Darkness = factor(Darkness)) %>%
                           mutate(Flow = fct_relevel(Flow, c("Still Water","Flowing Water (2 BL/s)"))) %>%
                           mutate(Ablation = fct_relevel(Ablation, c("No Ablation","Ablated"))) %>%
                           mutate(Darkness = fct_relevel(Darkness, c("Light","Dark"))) %>%
                           mutate(Flow_Ablation_Darkness = factor(paste(Flow,Ablation,Darkness,sep=", "))) %>%
                           separate(Fish, c("Fish1", "Fish2"),sep="x") %>% 
                           # ### Used to remove every third point to reduce autocorrelation
                           mutate(Data_ID = paste(Year,Month,Day,Trial, sep = "_"),
                                  Date = paste(Year,Month,Day,Trial, sep = "_"),
                                  Fish_ID = paste(Year,Month,Day,Trial,Fish1, sep = "_"),
                                  Comp_ID = paste(Year,Month,Day,Trial,Fish1,Fish2, sep = "_")) %>%
                           group_by(Comp_ID) %>%
                           filter(row_number() %% 3 == 1) %>%
                           ungroup()

@@

fish_comp_data <- rbind(fish_comp_data,single_comp_data)

min_dist <- 1
max_dist <- 3

#calculate average angle to all other fish in that range
fish_comp_data_ring_angles <- fish_comp_data %>% filter(Distance >= min_dist & Distance <= max_dist) %>%
                                                 group_by(Date,Year,Month,Day,Trial,
                                                          Flow,Ablation,Darkness,Flow_Ablation_Darkness,Fish1,Tailbeat_Num) %>%
                                                 summarise(Mean_Angle = rad_ang_mean(Angle),
                                                           #Should be the same number each time
                                                           Fish1_Yaw_Heading_PS = mean(Fish1_Yaw_Heading_PS, rm.na = T)) %>%
                                                 ungroup()

fish_data_nn_normal <- fish_comp_data_ring_angles %>% group_by(Flow_Ablation_Darkness) %>%
                                        mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS - mean(Fish1_Yaw_Heading_PS, na.rm = T)) %>%
                                        ungroup()

nn_heading_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Mean_Angle,pi/6)) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          se = sd(Fish1_Yaw_Heading_PS, na.rm = T) / sqrt(n()),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit() %>%
                                                group_by(Flow_Ablation_Darkness) %>%
                                                mutate(se = mean(se)) %>%
                                                ungroup()
```

Load in Single Fish

@@

heading_ps_angle_step <- stepAIC(heading_ps_angle_base,
                                   list(upper = ~ Mean_Angle*sin(Mean_Angle)*Ablation*Darkness, lower = ~1),
                                   direction = "both",
                                   trace = F)

@@

manual_model <- lmer(Fish1_Yaw_Heading_PS ~ sin(Mean_Angle) + sin(Mean_Angle):Ablation + sin(Mean_Angle):Darkness + sin(Mean_Angle):Ablation:Darkness + (1|Date), 
                   data = fish_data_nn_model) 
Anova(manual_model)

#Raw data
ggplot(fish_data_nn_model,
       aes(x = Mean_Angle, y = rad2deg(Fish1_Yaw_Heading_PS), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.5)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_line(aes(y = predict(manual_model)), size = 1)+
  theme_classic() +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color,"red")) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")

graph_model_data = nn_heading_sum %>% filter(Flow == "Still Water") %>% mutate(Mean_Angle = angle_bins)

#Means graph
ggplot(nn_heading_sum %>% filter(Flow == "Still Water"),
       aes(x = angle_bins, y = rad2deg(mean_turn), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  geom_line(aes( y = rad2deg(predict(manual_model,newdata = graph_model_data, re.form = NA)) ), size = 1)+
  geom_ribbon(aes(ymin = rad2deg(predict(manual_model,newdata = graph_model_data, re.form = NA) - se),
                  ymax = rad2deg(predict(manual_model,newdata = graph_model_data, re.form = NA) + se),
                  fill = interaction(Ablation, Darkness, sep =", ")),
              alpha = 0.25, colour = NA)+
  theme_classic() +
  ylim(-75,75) +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle speed (degrees/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color,"red")) +
  scale_fill_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color,"red")) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")+
  ggtitle(paste0("All Neighbors From ",min_dist," to ",max_dist))+ 
  guides(fill = FALSE)

test(emmeans(manual_model, trt.vs.ctrlk ~ Ablation*Darkness | Mean_Angle,
             at = list(Mean_Angle = c(pi/2))))

@@
